---
title: "Creating graphs in R with `ggplot2`"
author: "Jewel Uzquiano"
date: "10/11/2021"
output: html_document
---

# Plots with ggplot2

We are going to make some plots in R.

```{r}
library(tidyverse)
library(readxl)
library(here)
```
```{r}
ca_np <- read_csv(here("data", "ca_np.csv"))
names(ca_np)
head(ca_np)
summary(ca_np)

ci_np <- read_excel(here("data", "ci_np.xlsx"))
```

#Making the first ggplot graph
Visualizing visitors to Channel Islands National Park

```{r}
gg_base <- ggplot(data = ci_np,
      aes(x = year, y = visitors))

gg_base +
  geom_line()

gg_base +
  geom_point()

gg_base +
  geom_col()

gg_base +
  geom_area()
```

We can also start customizing `ggplot` graphs

```{r}
gg_base +
  geom_line(
    color = "purple",
    linetype = "dashed"
  )

gg_base +
  geom_point(
    color = "purple",
    pch = 17,
    size = 4,
    alpha = 0.5
  )

gg_base +
  geom_area(
    color = "green",
    fill = "green",
    alpha = 0.5
  )

## mapping variables onto aesthetics
gg_base +
  geom_point(
    aes(size = visitors,
        color = visitors),
    alpha = 0.5
  )

gg_base +
  geom_col(aes(fill = visitors))
```
#Learning about themes

```{r}
gg_base +
  geom_point(
    aes(size = visitors,
        color = visitors),
    alpha = 0.5
  ) +
  theme_minimal()

gg_base +
  geom_line(
    linetype = "dotted" 
  ) +
  theme_bw() +
  labs(
    x = "Year",
    y = "Annual park visitors",
    title = "Channel Islands NP Visitation",
    subtitle = "(1963 - 2016)"
  )
```
# We can also combine geoms

```{r}
gg_base +
  geom_line(color = "purple") +
  geom_point(color = "orange",
             aes(size = year),
             alpha = 0.5)
```
# multi series ggplot graphs

```{r}
ggplot(data = ca_np, aes(x = year, y = visitors, group = park_name)) +
  geom_line()

gg_np <- ggplot(data = ca_np, aes(x = year, y = visitors, group = park_name))

gg_np +
  geom_line(show.legend = FALSE) +
  theme_light() +
  labs(
    x = "year",
    y = "annual visitors") +
  facet_wrap(~ park_name)

```
We can use `ggsave` to export. ggsave() will export the most recently run ggplot graph by default (plot = last_plot()), unless you give it the name of a different saved ggplot object. Some common arguments for ggsave():

width =: set exported image width (default inches)
height =: set exported image height (default height)
dpi =: set dpi (dots per inch)

```{r}
ggsave(here("figures", "np_graph.jpg"), dpi = 180, width = 8, height = 7)

```

